A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-03-16 paper/theor-experiment.tex
more changes to awa section of paper

The first two objectives minimize the distance from the position of the current
  minimum peak to the expected peak location at $3.0$\,m for transverse bunch
  size (beam waist) and emittance (see Figure~\ref{fig:fer_match}).
The third objective (\ref{eq:swissfel:p3}) adds a condition preferring
  solutions that have their emittance and rms peak locations at the same
  $z$-coordinate.
Equations (\ref{eq:swissfel:firstconstr}) and (\ref{eq:swissfel:lastconstr})
  define constraints for initial conditions for the simulation: charge,
  gun voltage and laser spot size.
Design variables given in (\ref{eq:swissfel:firstdvar}) to
  (\ref{eq:swissfel:lastdvar}) correspond to field strengths of the
  first focusing magnet, its displacement, and the phase of the gun.

@@

For validation purposes we compared the results of the envelope-tracker using
  the analytical space charge model with the \textsc{OPAL} 3D macro particle
  tracker.
The benchmark was run on the first 12.5 meters of the \textsc{SwissFEL} 250
  MeV injector.
The results for both rms beamsize and emittance are shown in
  Figure~\ref{fig:et-vs-tt}.
A good agreement between the two codes can be observed.
The difference of the larger emittance along the solenoids in case of 3D
  tracker that is not seen by the envelope-tracker is due to the different
  definition of the particle momenta (canonical vs.~mechanical).
Both trackers agree within acceptable limits \cite{chao:99}.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{Report/iff_plot_emrms_ettt}
  \caption{Comparison 3D-tracker versus envelope-tracker in case for rms$_{x}$
    and $\varepsilon_{x}$.}
  \label{fig:et-vs-tt}
\end{figure}

@@

\subsection{AWA Photoinjector Optimization}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles. This beam line has served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [XX], and will be used in the future for 
similar experiments. One of the main constraints on prior experimental 
results was the beam size and emittance as the beam passed through small aperture 
wakefield structures. In an attempt to maximize charge transmission in future experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.

Compared to the Ferrario experiment in Section \ref{ferrario}, 
this problem is larger by a factor of two in both design variables and objectives.
The problem is also less predictable, meaning there is not a clear idea
before hand where the pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.

The design variables include both solenoid current, IBF and IM,
the gun phase, GPHASE, and the laser full width half maximum in 
the longitudinal direction. The definition of the design variables 
was included in the OPAL input file, 
As in Section \ref{ferrario}, 
we also have to define several beam line parameters that 
are not included in the optimization design variables above.
These are shown in Table \ref{awatable}.

\begin{align}
\text{min}  \quad & \text{rms}_{x, s1}, \quad \text{rms}_{x, s2} \label{eq:awa:p1}\\
& \text{rms}_{s, s1}, \quad \text{rms}_{s, s2} \label{eq:awa:p2}\\
& \varepsilon_{x,s1}, \quad \varepsilon_{x,s2} \label{eq:awa:p3} \\
\text{subject to} \quad & q = 200 \left[\text{pC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{RF}} = 100 \left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \sigma_{L} \leq \sigma_x = \sigma_y \leq \sigma_{U} \label{eq:awa:firstdvar}\\
\quad & \text{KS}_{L} \leq \text{KS}_{\text{RF}} \leq \text{KS}_{U} \label{eq:awa:seconddvar}\\
\quad & \text{LAG}_{L} \leq \text{LAG}_{\text{RF}} \leq \text{LAG}_{U} \\
\quad & \Delta z_{L\text{KS}} \leq \Delta z_{\text{KS}} \leq \Delta z_{U\text{KS}} \label{eq:awa:lastdvar}
\end{align}

>>>>

The first two objectives minimize the distance from the position of the current
  minimum peak to the expected peak location at $3.0$\,m for transverse bunch
  size (beam waist) and emittance (see Figure~\ref{fig:fer_match}).
The third objective (\ref{eq:swissfel:p3}) adds a condition preferring
  solutions that have their emittance and rms peak locations at the same
  $s$-coordinate.
Equations (\ref{eq:swissfel:firstconstr}) and (\ref{eq:swissfel:firstdvar})
  define constraints for initial conditions for the simulation: charge,
  gun voltage and laser spot size.
Design variables given in (\ref{eq:swissfel:seconddvar}) to
  (\ref{eq:swissfel:lastdvar}) correspond to field strengths of the
  first focusing magnet, its displacement, and the phase of the gun.

@@

For validation purposes we compared the results of the envelope-tracker using
  the analytical space charge model with the \textsc{OPAL} 3D macro particle
  tracker.
The benchmark was run on the first 12.5 meters of the \textsc{SwissFEL} 250
  MeV injector.
The results for both rms beamsize and emittance are shown in
  Figure~\ref{fig:et-vs-tt}.
A good agreement between the two codes can be observed.
The difference of the larger emittance along the solenoids in case of 3D
  tracker that is not seen by the envelope-tracker is due to the different
  definition of the particle momenta (canonical vs.~mechanical).
Both trackers agree within acceptable limits \cite{chao:99}.
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{Report/iff_plot_emrms_ettt}
  \caption{Comparison 3D-tracker versus envelope-tracker in case for rms$_{x}$
    and $\varepsilon_{x}$.}
  \label{fig:et-vs-tt}
\end{figure}

@@

\subsection{AWA Photoinjector Optimization}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles. This beam line has served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [XX], and will be used in the future for 
similar experiments. One of the main constraints on prior experimental 
results was the beam size and emittance as the beam passed through small aperture 
wakefield structures. In an attempt to maximize charge transmission in future experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.
\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Report/awa-drawing}
	\caption{High charge beam line layout at the AWA.}
	\label{awa-pic}
\end{figure}

Compared to the Ferrario experiment in Section \ref{ferrario}, 
we increase the number of design variables and objectives by a factor of two.
The problem is also less predictable, due to the large number of design variables
and nonlinear effects such as space charge, there is not a definite idea
before hand where the Pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.

We define the AWA optimization problem using objectives, and parameters given in 
equations (\ref{eq:awa:p1}) to (\ref{eq:awa:lastdvar}).

\begin{align}
\text{min}  \quad & \text{rms}_{x, s1}, \quad \text{rms}_{x, s2} \label{eq:awa:p1}\\
& \text{rms}_{s, s1}, \quad \text{rms}_{s, s2} \label{eq:awa:p2}\\
& \varepsilon_{x,s1}, \quad \varepsilon_{x,s2} \label{eq:awa:p3} \\
\text{subject to} \quad & q = 40 \left[\text{nC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64\left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24-25\left[\text{MV/m}\right] \\
\quad & \sigma_{L} \leq \sigma_x = \sigma_y \leq \sigma_{U} \label{eq:awa:firstdvar}\\
\quad & \text{KS}_{L} \leq \text{KS}_{\text{RF}} \leq \text{KS}_{U} \label{eq:awa:seconddvar}\\
\quad & \text{LAG}_{L} \leq \text{LAG}_{\text{RF}} \leq \text{LAG}_{U} \\
\quad & \Delta z_{L\text{KS}} \leq \Delta z_{\text{KS}} \leq \Delta z_{U\text{KS}} \label{eq:awa:lastdvar}
\end{align}

The first four objectives minimize the transverse ($rms_x$) and longitudinal ($rms_s$) 
beam sizes at to locations of interest in the beam line, [$s_1$,$s_2$]. 
These are the longitudinal positions of 
the entrance of the kicker and exit of the kicker. The beam size needs to be minimized
here to prevent scraping and ensure transmission through the wakefield structures 
following the kicker. The last two objectives (\ref{eq:awa:p3}) minimize the 
transverse emittance, which favors less divergence in the beam.
